---
title: "Team-4-Case-One-Study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyverse)
library(GGally)
library(plotly)
library(ggplot2)
library(plyr) # Load first before dplyer 
library(dplyr)
library(tidyr)
library(maps)
library(mapproj)
library(ggalt)
library(class)
library(caret)
library(e1071)
library(tm)
library(mvtnorm)
library(xlsx)
library(leaflet)
library(ggvis)
```

## Setup
```{r}
# Setup file locations to import 
beer_csv_loc <- "./StartingDBs/Beers.csv"
breweries_csv_loc <- "./StartingDBs/Breweries.csv"

# Read in data 
beer_data <- read.csv(beer_csv_loc, header = TRUE)
brewery_data <- read.csv(breweries_csv_loc, header = TRUE)

# Data provided by Celia
# Celia's contributions prefied by cb_
cb_beer <- read.xlsx(file="./StartingDBs/Beers_Celia.xlsx", sheetIndex = 1, 
                     header=TRUE, stringsAsFactors = FALSE)
cb_brewery <- read.xlsx(file="./StartingDBs/Breweries_Celia.xlsx", sheetIndex = 1, 
                     header=TRUE, stringsAsFactors = FALSE)
```


# Case study Questions 

## 1. How many breweries are present in each state?
```{r}
# Create a seperate data frame to store state data 
### Note: This data set counts the district of columbia as a state
num_of_breweries_by_state <- data.frame(table(brewery_data$State))

##### Celia Banks #####
#Begin prepare an interactive geo map of breweries by state
#set the map
leaflet() %>% addTiles()
#add popup for brewery information
brewery_map <- cb_brewery %>% mutate(popup_info=paste(Name,"<br/>",City,"<br>",State,"<br>"))
#tweak the map with info layout, set circles smaller to enhance visual
leaflet() %>% addTiles() %>% addCircleMarkers(data=brewery_map, 
                                              lat=~Latitude, 
                                              lng=~Longitude, 
                                              radius=~1, 
                                              popup=~popup_info)
#End prepare an interactive geo map of breweries by state#

#corresponding distribution plot of breweries by state
ggplot(brewery_map, aes(x=as.factor(State), color="blue")) +
       geom_bar(color="blue") +
       ggtitle("Distribution of Breweries by State") +
       xlab("State") +
       geom_text(stat='count', aes(label=after_stat(count)), vjust=-.25) +
       theme(legend.position="none", axis.text.x = element_text(angle = 90)) +
       theme(plot.title = element_text(hjust = 0.5))

#corresponding distribution plot of breweries by region
ggplot(brewery_map, aes(x=as.factor(Region), color="green")) +
       geom_bar(color="green") +
       ggtitle("Distribution of Breweries by Region") +
       xlab("Region") +
       geom_text(stat='count', aes(label=after_stat(count)), vjust=-.25) +
       theme(legend.position="none", axis.text.x = element_text(angle = 90)) +
       theme(plot.title = element_text(hjust = 0.5))

##### END CB #####
```


## 2. Merge beer data with the breweries data. Print the first 6 observations and
#the last six observations to check the merged file
```{r}
# Change column name of beer_data to match brewery_data to use as a primary key
colnames(beer_data)[5] <- "Brew_ID"

# Merge two data bases using merge
full_brew_data <- merge(brewery_data, beer_data, by="Brew_ID")

# Rename columns 2 and 5 that were changed during the merge
colnames(full_brew_data)[2] <- "Brewery"
colnames(full_brew_data)[5] <- "Beer_Name"

# Print first 6 and last 6 observations 
head(full_brew_data, n=6)
tail(full_brew_data, n=6)

# Create CSV Files for head and tail 
head_output_csv <- "./output/head_beer_data.csv"
tail_output_csv <- "./output/tail_been_data.csv"
write.csv(head(full_brew_data, n=6), head_output_csv)
write.csv(tail(full_brew_data, n=6), tail_output_csv)
```


## 3. Address any <NA> values
```{r}
# Create two new tables where one has only ABV with no <NA>'s and the other 
state_abv <- data.frame(State=full_brew_data$State, ABV=full_brew_data$ABV)
state_abv <- state_abv %>% drop_na(ABV)

# Do the same for IBU 
state_ibu <- data.frame(State=full_brew_data$State, IBU=full_brew_data$IBU)
state_ibu <- state_ibu %>% drop_na(IBU)
```


## 4. Compute the median alcohol content and international bitterness unit for each
##state. Plot a bar chart 
```{r, fig.height=7, fig.width=12}
# Find the average ABV for the state_abv table 
avg_state_abv <- ddply(state_abv, .(State), function(x) median(x$ABV))
colnames(avg_state_abv)[2] <- "Average_ABV" # Rename the column
# Drop DC. It's not a state....yet
avg_state_abv <- avg_state_abv[-c(8),] # DC is equal to position 8 on the table 
# Order the states by Decending ABV content 
avg_state_abv <- avg_state_abv[order(avg_state_abv$Average_ABV, decreasing=TRUE),]

# Find Average IBU
avg_state_ibu <- ddply(state_ibu, .(State), function(x) median(x$IBU))
colnames(avg_state_ibu)[2] <- "Average_IBU" # Rename column
# Order the states by IBU Decending 
avg_state_ibu <- avg_state_ibu[order(avg_state_ibu$Average_IBU, decreasing=TRUE),]

# Plot Average Alcohol by Volume per state
ggplot(avg_state_abv, aes(x=reorder(State, desc(Average_ABV)), y=Average_ABV, fill=State)) + 
       geom_col(show.legend = FALSE, width=.9, position="dodge") +
       ggtitle("Average ABV per State") +
       xlab("State") + 
       ylab("Average Alcohol by Volume") 

# Plot Average International Bitterness Unit per State 
ggplot(avg_state_ibu, aes(x=reorder(State, desc(Average_IBU)), y=Average_IBU, fill=State)) + 
       geom_col(show.legend = FALSE, width=.9, position="dodge") +
       ggtitle("Average IBU per State") + 
       xlab("State") + 
       ylab("Average IBU")
```


## 4.1 Additional data provided by Celia 
```{r}
##### Celia Code #####
merge_bb <- merge(cb_beer, cb_brewery, by.x="Brewery_id", by.y="Brew_ID")
#Drop latitude and longitude columns as no longer needed
drop <- c("Latitude","Longitude")

cb_merged_bb <- merge_bb[,!(names(merge_bb) %in% drop)]

#rename columns to avoid confusion
#cb_merged_bb <- cb_merged_bb %>% rename(Beer_Name = Name.x, Brewery_Name = Name.y)
### RAH ###
colnames(cb_merged_bb)[2] <- "Beer_Name"
colnames(cb_merged_bb)[8] <- "Brewery_Name"
### End RAH ###

#get count of breweries in each state
brewery_count <- aggregate(Brewery_id ~ State, cb_merged_bb, sum)
aggregate(cb_merged_bb$Brewery_id, by=list(Category=cb_merged_bb$State), FUN=sum)

#impute the rows with zeroes
cb_merged_bb$ABV[is.na(cb_merged_bb$ABV)] <- 0
cb_merged_bb$IBU[is.na(cb_merged_bb$IBU)] <- 0

#compute median ABV and IBU for each state
cb_medianABV <- aggregate(ABV ~ State + Region, cb_merged_bb, median)
cb_medianIBU <- aggregate(IBU ~ State + Region, cb_merged_bb, median)

#merge the medians dataframes
cb_mergeABVIBU <- merge(cb_medianABV, cb_medianIBU, by="State")
drop <- c("Region.y")
cb_mergeABVIBU = merge_bb[,!(names(merge_bb) %in% drop)]

#Visualize ABV v IBU medians by state and by region
options(scipen = 999)
cb_mergeABVIBU <- as.data.frame(cb_mergeABVIBU)

#by state
ggplot(cb_mergeABVIBU, aes(ABV, IBU, color=State)) +
       geom_point(show.legend = FALSE) +
       facet_wrap(vars(State)) 

#by region
ggplot(cb_mergeABVIBU, aes(ABV, IBU, color=Region)) +
       geom_point() +
       facet_wrap(vars(Region))
##### END #####
```



## 5. Which state has the the Max ABV. Which has the 
```{r}
# Get the State with the highest average ABV
state_max_abv <- avg_state_abv[which.max(avg_state_abv$Average_ABV),] # NV
# Get the State with the highest average IBU
state_max_ibu <- avg_state_ibu[which.max(avg_state_ibu$Average_IBU),] # WV
```
### Nevada has the highest percentage of alcohal by volume while West Virginia has the highest IBU. 


## 6. Comment on the Summary stats and distribution of the ABV variable 
```{r}
summary(state_abv)
ggplot(state_abv, aes(x=ABV)) + geom_histogram() # Right Skewness
ggplot(state_abv, aes(x=ABV)) + geom_boxplot()
```


##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? 
```{r}
# Create a new data frame from the beer data that removes rows that have <NA> in 
# the ABV and IBU columns
beer_data_no_na <- beer_data
beer_data_no_na <- beer_data_no_na %>% drop_na(ABV)
beer_data_no_na <- beer_data_no_na %>% drop_na(IBU)

# Scatter plot to show raw data 
ggplot(beer_data_no_na, aes(x=ABV, y=IBU, colour=Style)) + 
       geom_point(show.legend = FALSE) + 
       ggtitle("ABV vs IBU") +
       xlab("ABV in %") + 
       ylab("IBU")

# Plot a smooth curve to see a more "linear" pattern 
ggplot(beer_data_no_na, aes(x=ABV, y=IBU)) + 
       geom_smooth() +
       ggtitle("ABV vs IBU") + 
       xlab("ABV in %") + 
       ylab("IBU")
```


## 8. investigate the difference with respect to IBU and ABV between IPAs (India Pale Ales) 
#and other types of Ale. Use KNN.
```{r, include=FALSE}
# Create a data frame with only IPA's
#df1[grep("dog",df1$type),]
ipa_beers <- beer_data_no_na[grep("IPA",beer_data_no_na$Style),]
# Change the Sytle column to only have the value of IPA for easier model training
ipa_beers$Style <- "IPA"

ales_beer <- beer_data_no_na[grep("Ale",beer_data_no_na$Style),]
# Change the Sytle column to only have the value of Ale for easier model training
ales_beer$Style <- "Ale"

# Combine the two data sets 
ipas_and_ales <- rbind(ipa_beers, ales_beer)

# plot the data to get an initial view of the data
ggplot(ipas_and_ales, aes(x=ABV, y=IBU, colour=Style)) + geom_point()

# Create train and test data frames
ipas_and_ales_train <- ipas_and_ales # train data needs to be a duplicate of orig
ipas_and_ales_test <- data.frame()

# Split 70/30 (Train/Test)
samples_to_take <- round(as.numeric(nrow(ipas_and_ales_train)*.3), digits=0)
for (row in 1:samples_to_take) {
  tmp_row <- sample_n(ipas_and_ales_train, 1) # Take sample
  # Add Sample to test data
  ipas_and_ales_test <- ipas_and_ales_test %>% rbind(tmp_row)
  # Remove Sample from train data
  ipas_and_ales_train <- ipas_and_ales_train[!(ipas_and_ales_train$Beer_ID==tmp_row$Beer_ID),]
}

# find an optimal K value for the KNN model
style_accuracy <- data.frame(accuract=numeric(50), k=numeric(50))
for (iter in 1:50) {
  style_class <- knn(ipas_and_ales_train[,c(3,4)], ipas_and_ales_test[c(3,4)],
                     ipas_and_ales_train$Style,
                     prob=TRUE, k=iter)
  table(ipas_and_ales_test$Style, style_class)
  cm <- confusionMatrix(table(ipas_and_ales_test$Style, style_class))
  style_accuracy$accuracy[iter] <- cm$overall[1]
  style_accuracy$k[iter] <- iter
}
#plot(style_accuracy$k, style_accuracy$accuracy, type="l", xlab="k")
plot_ly(style_accuracy, x=style_accuracy$k, y=style_accuracy$accuracy, 
        type="scatter", mode="lines")

##
## On average this loop gives me an optimal k level of k=5,6,7
##


```


















