---
title: "Team-4-Case-One-Study"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(tidyverse)
library(GGally)
library(plotly)
library(ggplot2)
library(plyr) # Load first before dplyer 
library(dplyr)
library(tidyr)
library(maps)
library(mapproj)
library(ggalt)
library(class)
library(caret)
library(e1071)
library(tm)
library(mvtnorm)
```

## Setup
```{r}
# Setup file locations to import 
beer_csv_loc <- "./StartingDBs/Beers.csv"
breweries_csv_loc <- "./StartingDBs/Breweries.csv"

# Read in data 
beer_data <- read.csv(beer_csv_loc, header = TRUE)
brewery_data <- read.csv(breweries_csv_loc, header = TRUE)
```

# Case study Questions 

## 1. How many breweries are present in each state?
```{r}
# Create a seperate data frame to store state data 
### Note: This data set counts the district of columbia as a state
num_of_breweries_by_state <- data.frame(table(brewery_data$State))
```

## 2. Merge beer data with the breweries data. Print the first 6 observations and
#the last six observations to check the merged file
```{r}
# Change column name of beer_data to match brewery_data to use as a primary key
colnames(beer_data)[5] <- "Brew_ID"

# Merge two data bases using merge
full_brew_data <- merge(brewery_data, beer_data, by="Brew_ID")

# Rename columns 2 and 5 that were changed during the merge
colnames(full_brew_data)[2] <- "Brewery"
colnames(full_brew_data)[5] <- "Beer_Name"

# Print first 6 and last 6 observations 
head(full_brew_data, n=6)
tail(full_brew_data, n=6)
```

## 3. Address any <NA> values
```{r}
# Create two new tables where one has only ABV with no <NA>'s and the other 
state_abv <- data.frame(State=full_brew_data$State, ABV=full_brew_data$ABV)
state_abv <- state_abv %>% drop_na(ABV)

# Do the same for IBU 
state_ibu <- data.frame(State=full_brew_data$State, IBU=full_brew_data$IBU)
state_ibu <- state_ibu %>% drop_na(IBU)
```

## 4. Compute the median alcohol content and international bitterness unit for each
##state. Plot a bar chart 
```{r, fig.height=7, fig.width=12}
# Find the average ABV for the state_abv table 
avg_state_abv <- ddply(state_abv, .(State), function(x) mean(x$ABV))
colnames(avg_state_abv)[2] <- "Average_ABV" # Rename the column
# Drop DC. It's not a state....yet
avg_state_abv <- avg_state_abv[-c(8),] # DC is equal to position 8 on the table 
# Order the states by Decending ABV content 
avg_state_abv <- avg_state_abv[order(avg_state_abv$Average_ABV, decreasing=TRUE),]

# Find Average IBU
avg_state_ibu <- ddply(state_ibu, .(State), function(x) mean(x$IBU))
colnames(avg_state_ibu)[2] <- "Average_IBU" # Rename column
# Order the states by IBU Decending 
avg_state_ibu <- avg_state_ibu[order(avg_state_ibu$Average_IBU, decreasing=TRUE),]

# Plot Average Alcohol by Volume per state
ggplot(avg_state_abv, aes(x=reorder(State, desc(Average_ABV)), y=Average_ABV, fill=State)) + 
       geom_col(show.legend = FALSE, width=.9, position="dodge") +
       xlab("State") + 
       ylab("Average Alcohol by Volume") 

# Plot Average International Bitterness Unit per State 
ggplot(avg_state_ibu, aes(x=reorder(State, desc(Average_IBU)), y=Average_IBU, fill=State)) + 
       geom_col(show.legend = FALSE, width=.9, position="dodge") +
       xlab("State") + 
       ylab("Average IBU")
```

## 5. Which state has the the Max ABV. Which has the 
```{r}
# Get the State with the highest average ABV
state_max_abv <- avg_state_abv[which.max(avg_state_abv$Average_ABV),]
# Get the State with the highest average IBU
state_max_ibu <- avg_state_ibu[which.max(avg_state_ibu$Average_IBU),]
```

## 6. Comment on the Summary stats and distribution of the ABV variable 
```{r}
summary(state_abv)
ggplot(state_abv, aes(x=ABV)) + geom_histogram() # Right Skewness
```

##7. Is there an apparent relationship between the bitterness of the beer and its alcoholic content? 
```{r}
# Create a new data frame from the beer data that removes rows that have <NA> in 
# the ABV and IBU columns
beer_data_no_na <- beer_data
beer_data_no_na <- beer_data_no_na %>% drop_na(ABV)
beer_data_no_na <- beer_data_no_na %>% drop_na(IBU)

# Scatter plot to show raw data 
ggplot(beer_data_no_na, aes(x=ABV, y=IBU, colour=Style)) + 
                        geom_point(show.legend = FALSE) + 
                        xlab("ABV in %") + 
                        ylab("IBU")

# Plot a smooth curve to see a more "linear" pattern 
ggplot(beer_data_no_na, aes(x=ABV, y=IBU)) + 
                        geom_smooth() +
                        xlab("ABV in %") + 
                        ylab("IBU")
```



















